---
title: "Credit status classification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Credit status classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE,
  message  = FALSE
)
```

The dataset used is `credit_data` from the **modeldata** package. It contains
4,039 observations where the response variable `Status` is a binary
classification label ("good" or "bad"), predicted using 13 other covariates.

```{r data}
library(mlstudy)
library(modeldata)
library(yardstick)

data(credit_data, package = "modeldata")
credit_data <- tidyr::drop_na(credit_data)
```

## Training / test split

Split the dataset into training and test sets, then summarise the class
distribution in each set.

```{r split}
split <- ml_split(credit_data, target = "Status", prop_train = 0.80, seed = 3)

rsample::training(split) |>
  dplyr::count(Status) |>
  dplyr::mutate(prop = n / sum(n))

rsample::testing(split) |>
  dplyr::count(Status) |>
  dplyr::mutate(prop = n / sum(n))
```

## Model comparison via grid search

Three classification model families are tuned using cross-validated grid search:

- **Random Forest** — `mtry` and `min_n`
- **Elastic Net** (logistic regression) — `penalty` and `mixture`
- **XGBoost** — `tree_depth`, `learn_rate`, and `min_n`

A preprocessing recipe dummy-encodes nominal predictors, removes zero-variance
columns, and normalises numeric predictors before fitting any model.

`ml_compare()` runs the full pipeline: recipe → workflows → 5-fold CV grid
search → selects the winner by ROC AUC → final fit on the full training set →
evaluation on the held-out test set.

```{r compare, cache=TRUE}
result <- ml_compare(
  credit_data,
  target      = "Status",
  prop_train  = 0.80,
  seed_split  = 3,
  seed_tune   = 42,
  cv_folds    = 5,
  rf_trees    = 300,
  xgb_trees   = 300,
  grid_levels = 4
)
```

## Cross-validated model comparison

```{r cv-compare}
result$cv_compare[, c("model", "mean", "std_err", "n")]
```

Best hyper-parameters per model:

```{r best-params}
result$best_params
```

## Tuning plot for the winning model

The plot below mirrors the tuning visualisation from self-study 1.

```{r tune-plot, fig.width=7, fig.height=4}
if (result$best == "XGBoost") {
  tune::collect_metrics(result$tune_results[["XGBoost"]]) |>
    dplyr::filter(.metric == "roc_auc") |>
    dplyr::mutate(
      tree_depth = factor(tree_depth),
      min_n      = factor(min_n)
    ) |>
    ggplot2::ggplot(ggplot2::aes(
      x      = learn_rate,
      y      = mean,
      colour = tree_depth,
      group  = tree_depth
    )) +
    ggplot2::geom_line(linewidth = 1.1, alpha = 0.7) +
    ggplot2::geom_point(size = 2) +
    ggplot2::facet_wrap(~min_n) +
    ggplot2::labs(
      title    = "XGBoost tuning results",
      subtitle = "Cross-validated ROC AUC across learn_rate, tree_depth, and min_n",
      x        = "learn_rate",
      y        = "Mean ROC AUC",
      colour   = "tree_depth"
    )
}
```

## Test-set evaluation

```{r test-metrics}
tune::collect_metrics(result$final_res)
```

Individual metrics on the test set:

```{r individual-metrics}
preds <- result$test_preds

roc_auc(preds,   truth = Status, .pred_bad)
accuracy(preds,  truth = Status, estimate = .pred_class)
precision(preds, truth = Status, estimate = .pred_class, event_level = "second")
```

ROC curve:

```{r roc-curve, fig.width=5, fig.height=5}
preds |>
  roc_curve(truth = Status, .pred_bad) |>
  ggplot2::autoplot()
```

Confusion matrix:

```{r conf-mat}
conf_mat(preds, truth = Status, estimate = .pred_class)
```

## Package print / summary / plot methods

```{r print}
print(result)
```

```{r summary}
summary(result)
```

```{r plot, fig.width=7, fig.height=4}
plot(result)
```
